### 2015-12-08 ada-ftrl.tex
Many improvements
1) Add conclusion
2) Add limits and their explanations
3) Add many bibliographic references
4) Fix formatting for most .tex files

The sequence $\{\Delta_t\}_{t=0}^\infty$ is non-negative and non-decreasing.
Also, $\Delta_t$ as a function of $\ell_1, \ell_2, \dots, \ell_t$ is positive homogenous
of degree one, making the algorithm scale-free.

If $\Delta_{i-1} = 0$, we define $\Delta_{i-1}
\Breg_{R^*}(\frac{-L_i}{\Delta_{i-1}}, \frac{-L_{i-1}}{\Delta_{i-1}})$ as
$\lim_{a \to 0^+} a \Breg_{R^*}(\frac{-L_i}{a}, \frac{-L_{i-1}}{a})$ which
always exists and is finite; see Lemma~\ref{lemma:limit-bregman-divergence}
in~\ref{section:limits}.  Similarly, when $\Delta_{t-1} = 0$, we define $w_t =
\argmin_{w \in K} \langle L_{t-1}, w \rangle$ where ties among minimizers are
broken by taking the one with the smallest value of $R(w)$, which is unique due
to strong convexity; as we show in Lemma~\ref{lemma:prediction-limit-existence}
in~\ref{section:limits}, this is the same as $w_t = \lim_{a \to 0^+}
\argmin_{w \in K} (\langle L_{t-1}, w \rangle + aR(w))$.

@@

Since, $\Breg_{R_t^*}(u,v) = \Delta_{t-1} \Breg_{R^*}(\frac{u}{\Delta_{t-1}},
\frac{v}{\Delta_{t-1}})$ by definition of Bregman divergence and Part 8 of
Proposition~\ref{proposition:conjugate-properties}, we have $\sum_{t=1}^T
\Breg_{R_t^*}(-L_t, -L_{t-1}) = \Delta_T$.
\end{proof}

>>>>

The sequence $\{\Delta_t\}_{t=0}^\infty$ is non-negative and non-decreasing.
Also, $\Delta_t$ as a function of $\ell_1, \ell_2, \dots, \ell_t$ is positive
homogeneous of degree one, making the algorithm scale-free.

If $\Delta_{i-1} = 0$, we define $\Delta_{i-1}
\Breg_{R^*}(\frac{-L_i}{\Delta_{i-1}}, \frac{-L_{i-1}}{\Delta_{i-1}})$ as
$\lim_{a \to 0^+} a \Breg_{R^*}(\frac{-L_i}{a}, \frac{-L_{i-1}}{a})$ which
always exists and is finite; see Lemma~\ref{lemma:limit-bregman-divergence}
in~\ref{section:limits}.  Similarly, when $\Delta_{t-1} = 0$, we define $w_t =
\argmin_{w \in K} \langle L_{t-1}, w \rangle$ where ties among minimizers are
broken by taking the one with the smallest value of $R(w)$, which is unique due
to strong convexity. As we show in Lemma~\ref{lemma:prediction-limit-existence}
in~\ref{section:limits}, this is the same as $w_t = \lim_{a \to 0^+} \argmin_{w
\in K} (\langle L_{t-1}, w \rangle + aR(w))$.

@@

Since, $\Breg_{R_t^*}(u,v) = \Delta_{t-1} \Breg_{R^*}(\frac{u}{\Delta_{t-1}},
\frac{v}{\Delta_{t-1}})$ by definition of Bregman divergence and property 8 of
Proposition~\ref{proposition:conjugate-properties}, we have $\sum_{t=1}^T
\Breg_{R_t^*}(-L_t, -L_{t-1}) = \Delta_T$.
\end{proof}

### 2015-12-08 introduction.tex
Many improvements
1) Add conclusion
2) Add limits and their explanations
3) Add many bibliographic references
4) Fix formatting for most .tex files

OLO is a fundamental problem in machine
learning~\cite{Cesa-Bianchi-Lugosi-2006, Rakhlin-Sridharan-2009,
Shalev-Shwartz-2011}.  Many learning problems can be directly phrased as OLO,
e.g., learning with expert advice~\cite{Littlestone-Warmuth-1994, Vovk-1998,
Freund-Schapire-1997, Cesa-Bianchi-Haussler-Helmbold-Schapire-Warmuth-1997},
online combinatorial optimization~\cite{Kalai-Vempala-2005,
Helmbold-Warmuth-2009, Koolen-Warmuth-Kivinen-2010}. Other problems can be
reduced to OLO, e.g. online convex
optimization~\cite[Chapter~2]{Shalev-Shwartz-2011}, online classification
~\cite{Rosenblatt-1958,Freund-Schapire-1999} and
regression~\cite{Kivinen-Warmuth-1997},
~\cite[Chapters~11~and~12]{Cesa-Bianchi-Lugosi-2006}, multi-armed
problems~\cite[Chapter~6]{Cesa-Bianchi-Lugosi-2006},
\cite{Bubeck-Cesa-Bianchi-2012}, and batch and stochastic optimization of
convex functions~\cite{Nemirovski-Yudin-1983, Bubeck-2015}.  Hence, a result in
OLO immediately implies other results in all these domains.

@@

\begin{table}[t]
\fontsize{8}{8.2}\selectfont
\centering
\begin{tabular}{|p{3.6cm}|c|p{3.4cm}|c|}
\hline
\textbf{Algorithm} & \textbf{Decisions Set(s)} & \textbf{Regularizer(s)} & \textbf{Scale-Free} \\ \hline \hline
\textsc{Hedge} \cite{Freund-Schapire-1997} & Probability Simplex & Negative Entropy & No \\ \hline
\textsc{GIGA} \cite{Zinkevich-2003} & Any Bounded & $\frac{1}{2}\|w\|_2^2$ & No \\ \hline
\textsc{RDA} \cite{Xiao-2010} & \textbf{Any} & \textbf{Any Strongly Convex} & No \\ \hline
\textsc{FTRL-Proximal} \cite{McMahan-Streeter-2010,McMahan-2014} & Any Bounded & $\frac{1}{2}\|w\|_2^2 + $ any convex func. & \textbf{Yes} \\ \hline
\textsc{AdaGrad MD} \cite{Duchi-Hazan-Singer-2011} & Any Bounded & $\frac{1}{2}\|w\|_2^2 + $ any convex func. & \textbf{Yes} \\ \hline
\textsc{AdaGrad FTRL} \cite{Duchi-Hazan-Singer-2011} & \textbf{Any} & $\frac{1}{2}\|w\|_2^2 + $ any convex func. & No \\ \hline
\textsc{AdaHedge} \cite{de-Rooij-van-Erven-Grunwald-Koolen-2014} & Probability Simplex & Negative Entropy & \textbf{Yes} \\ \hline
\textsc{Optimistic MD} \cite{Rakhlin-Sridharan-2013} & $\sup_{u,v \in K} \Breg_f(u,v) < \infty$ & \textbf{Any Strongly Convex} & \textbf{Yes} \\ \hline
\textsc{NAG} \cite{Ross-Mineiro-Langford-2013} & $\{u: \max_t \langle \ell_t, u\rangle \leq C\}$ & $\frac{1}{2}\|w\|_2^2 $& Partially\footnotemark \\ \hline
\textsc{Scale invariant algorithms} \cite{Orabona-Crammer-Cesa-Bianchi-2014} & \textbf{Any} & $\frac{1}{2}\|w\|_p^2 + $ any convex func. \newline $1 < p \le 2$ & Partially\textsuperscript{\ref{footnote-label}} \\ \hline
\textsc{AdaFTRL} \textbf{[this paper]} & Any Bounded & \textbf{Any Strongly Convex} & \textbf{Yes} \\ \hline
\textsc{SOLO FTRL} \textbf{[this paper]} & \textbf{Any} & \textbf{Any Strongly Convex} & \textbf{Yes} \\ \hline
\end{tabular}
\caption{Selected results for OLO. Best results in each column are in bold.
\label{table:results}
}
\end{table}

>>>>

OLO is a fundamental problem in machine
learning~\cite{Cesa-Bianchi-Lugosi-2006, Rakhlin-Sridharan-2009,
Shalev-Shwartz-2011}.  Many learning problems can be directly phrased as OLO,
e.g., learning with expert advice~\cite{Littlestone-Warmuth-1994, Vovk-1998,
Freund-Schapire-1997, Cesa-Bianchi-Haussler-Helmbold-Schapire-Warmuth-1997},
online combinatorial optimization~\cite{Kalai-Vempala-2005,
Helmbold-Warmuth-2009, Koolen-Warmuth-Kivinen-2010}. Other problems can be
reduced to OLO, e.g. online convex
optimization~\cite{Abernethy-Bartlett-Rakhlin-Tewari-2008},
\cite[Chapter~2]{Shalev-Shwartz-2011}, online classification
~\cite{Rosenblatt-1958,Freund-Schapire-1999} and
regression~\cite{Kivinen-Warmuth-1997},
~\cite[Chapters~11~and~12]{Cesa-Bianchi-Lugosi-2006}, multi-armed
problems~\cite[Chapter~6]{Cesa-Bianchi-Lugosi-2006},
\cite{Abernethy-Hazan-Rakhlin-2008, Bubeck-Cesa-Bianchi-2012}, and batch
and stochastic optimization of convex functions~\cite{Nemirovski-Yudin-1983,
Bubeck-2015}.  Hence, a result in OLO immediately implies other results in all
these domains.

@@

\begin{table}[t]
\fontsize{8}{8.2}\selectfont
\centering
\begin{tabular}{|p{3.6cm}|c|p{3.4cm}|c|}
\hline
\textbf{Algorithm} & \textbf{Decisions Set(s)} & \textbf{Regularizer(s)} & \textbf{Scale-Free} \\ \hline \hline
\textsc{Hedge} \cite{Freund-Schapire-1997} & Probability Simplex & Negative Entropy & No \\ \hline
\textsc{GIGA} \cite{Zinkevich-2003} & Any Bounded & $\frac{1}{2}\|w\|_2^2$ & No \\ \hline
\textsc{RDA} \cite{Xiao-2010} & \textbf{Any} & \textbf{Any Strongly Convex} & No \\ \hline
\textsc{FTRL-Proximal} \cite{McMahan-Streeter-2010,McMahan-2014} & Any Bounded & $\frac{1}{2}\|w\|_2^2 + $ any convex func. & \textbf{Yes} \\ \hline
\textsc{AdaGrad MD} \cite{Duchi-Hazan-Singer-2011} & Any Bounded & $\frac{1}{2}\|w\|_2^2 + $ any convex func. & \textbf{Yes} \\ \hline
\textsc{AdaGrad FTRL} \cite{Duchi-Hazan-Singer-2011} & \textbf{Any} & $\frac{1}{2}\|w\|_2^2 + $ any convex func. & No \\ \hline
\textsc{AdaHedge} \cite{de-Rooij-van-Erven-Grunwald-Koolen-2014} & Probability Simplex & Negative Entropy & \textbf{Yes} \\ \hline
\textsc{Optimistic MD} \cite{Rakhlin-Sridharan-2013} & $\sup_{u,v \in K} \Breg_f(u,v) < \infty$ & \textbf{Any Strongly Convex} & \textbf{Yes} \\ \hline
\textsc{NAG} \cite{Ross-Mineiro-Langford-2013} & $\{u: \max_t \langle \ell_t, u\rangle \le C\}$ & $\frac{1}{2}\|w\|_2^2 $& Partially\footnotemark \\ \hline
\textsc{Scale invariant algorithms} \cite{Orabona-Crammer-Cesa-Bianchi-2014} & \textbf{Any} & $\frac{1}{2}\|w\|_p^2 + $ any convex func. \newline $1 < p \le 2$ & Partially\textsuperscript{\ref{footnote-label}} \\ \hline
\textsc{AdaFTRL} \textbf{[this paper]} & Any Bounded & \textbf{Any Strongly Convex} & \textbf{Yes} \\ \hline
\textsc{SOLO FTRL} \textbf{[this paper]} & \textbf{Any} & \textbf{Any Strongly Convex} & \textbf{Yes} \\ \hline
\end{tabular}
\caption{Selected results for OLO. Best results in each column are in bold.
\label{table:results}
}
\end{table}

### 2015-12-08 limits.tex
Many improvements
1) Add conclusion
2) Add limits and their explanations
3) Add many bibliographic references
4) Fix formatting for most .tex files

Before we give the proof, we illustrate the lemma on a simple
example.  Let $K = [-1,1]^2$ be the unit square in $\R^2$ and let
$R(w) = \|w\|_2^2$. Let $L = (1,0)$. The minimizers are
$$
\argmin_{w \in K} \langle L, w \rangle = \{ (-1,y) ~:~ y \in [-1,1] \} \; .
$$
The minimizer with smallest value of $R(w)$ is $(-1,0)$. Hence the lemma implies
that
$$
\lim_{\eta \to +\infty} \argmin_{w \in K}
\left( \langle L, w \rangle + \frac{1}{\eta} \|w\|_2^2 \right) = (-1, 0) \; .
$$

@@

By non-negativity of $R$ and optimality of $w(\eta)$ we have
$$
\langle L, w(\eta) \rangle
\le \langle L, w(\eta) \rangle + \frac{1}{\eta} R(w(\eta))
\le \langle L, v^* \rangle + \frac{1}{\eta} R(v^*) \; .
$$
Taking the limit $\eta \to +\infty$, we see that
$$
\lim_{\eta \to +\infty} \langle L, w(\eta) \rangle \le
\lim_{\eta \to +\infty} \left( \langle L, v^* \rangle + \frac{1}{\eta} R(v^*) \right)
= \langle L, v^* \rangle \; .
$$
From~(\ref{equation:limit-is-minimizer}) we have
$\langle L, v^* \rangle \le \langle L, w \rangle$ for any $w$, and therefore
\begin{equation}
\label{equation:limit-equality}
\lim_{\eta \to +\infty} \langle L, w(\eta) \rangle = \langle L, v^* \rangle \; .
\end{equation}

@@

\begin{proof}
Using Part 3 of Proposition~\ref{proposition:conjugate-properties} we can write
the divergence
\begin{align*}
a \Breg_{R^*}(x/a, y/a) & = a R^*(x/a) - a R^*(y/a) - \langle x - y, \grad R^*(y/a) \rangle \\
& =
 a \left[ \langle x/a, \grad R^*(x/a) \rangle - R(\grad R^*(x/a)) \right] \\
& \qquad - a \left[ \langle y/a, \grad R^*(y/a) \rangle - R(\grad R^*(y/a)) \right]
- \langle x - y, \grad R^*(y/a) \rangle \\
& =
\langle x, \grad R^*(x/a) - \grad R^*(y/a) \rangle - a R(\grad R^*(x/a))
+ a R(\grad R^*(y/a)) \; .
\end{align*}
Part 2 of Proposition~\ref{proposition:conjugate-properties} implies that
\begin{align*}
u = \lim_{a \to 0^+} \grad R^*(x/a) & = \lim_{a \to 0^+} \argmin_{w \in K} \left( a R(w) - \langle x, w \rangle \right) \; , \\
v = \lim_{a \to 0^+} \grad R^*(y/a) & = \lim_{a \to 0^+} \argmin_{w \in K} \left( a R(w) - \langle y, w \rangle \right) \; .
\end{align*}
The limits on the right exist because of compactness of $K$. They are simply
the minimizers $u= \argmin_{w \in K} - \langle x, w \rangle$ and $v= \argmin_{w
\in K} - \langle y, w \rangle$ where ties in $\argmin$ are broken according to
smaller value of $R(w)$.

>>>>

Before we give the proof, we illustrate the lemma on a simple
example.  Let $K = [-1,1]^2$ be a closed square in $\R^2$ and let
$R(w) = \|w\|_2^2$. Let $L = (1,0)$. The minimizers are
$$
\argmin_{w \in K} \langle L, w \rangle = \{ (-1,y) ~:~ y \in [-1,1] \} \; .
$$
The minimizer with smallest value of $R(w)$ is $(-1,0)$. Hence the lemma implies
that
$$
\lim_{\eta \to +\infty} \argmin_{w \in K}
\left( \langle L, w \rangle + \frac{1}{\eta} \|w\|_2^2 \right) = (-1, 0) \; .
$$

@@

By non-negativity of $R$ and optimality of $w(\eta)$ we have
$$
\langle L, w(\eta) \rangle
\le \langle L, w(\eta) \rangle + \frac{1}{\eta} R(w(\eta))
\le \langle L, v^* \rangle + \frac{1}{\eta} R(v^*) \; .
$$
Taking the limit $\eta \to +\infty$, we see that
$$
\lim_{\eta \to +\infty} \langle L, w(\eta) \rangle
\le\lim_{\eta \to +\infty} \left( \langle L, v^* \rangle + \frac{1}{\eta} R(v^*) \right)
= \langle L, v^* \rangle \; .
$$
From~(\ref{equation:limit-is-minimizer}) we have
$\langle L, v^* \rangle \le \langle L, w \rangle$ for any $w$, and therefore
\begin{equation}
\label{equation:limit-equality}
\lim_{\eta \to +\infty} \langle L, w(\eta) \rangle = \langle L, v^* \rangle \; .
\end{equation}

@@

\begin{proof}
Using property 3 of Proposition~\ref{proposition:conjugate-properties} we can write
the divergence
\begin{align*}
a \Breg_{R^*}(x/a, y/a) & = a R^*(x/a) - a R^*(y/a) - \langle x - y, \grad R^*(y/a) \rangle \\
& =
 a \left[ \langle x/a, \grad R^*(x/a) \rangle - R(\grad R^*(x/a)) \right] \\
& \qquad - a \left[ \langle y/a, \grad R^*(y/a) \rangle - R(\grad R^*(y/a)) \right]
- \langle x - y, \grad R^*(y/a) \rangle \\
& =
\langle x, \grad R^*(x/a) - \grad R^*(y/a) \rangle - a R(\grad R^*(x/a))
+ a R(\grad R^*(y/a)) \; .
\end{align*}
Property 2 of Proposition~\ref{proposition:conjugate-properties} implies that
\begin{align*}
u = \lim_{a \to 0^+} \grad R^*(x/a) & = \lim_{a \to 0^+} \argmin_{w \in K} \left( a R(w) - \langle x, w \rangle \right) \; , \\
v = \lim_{a \to 0^+} \grad R^*(y/a) & = \lim_{a \to 0^+} \argmin_{w \in K} \left( a R(w) - \langle y, w \rangle \right) \; .
\end{align*}
The limits on the right exist because of compactness of $K$. They are simply
the minimizers $u= \argmin_{w \in K} - \langle x, w \rangle$ and $v= \argmin_{w
\in K} - \langle y, w \rangle$ where ties in $\argmin$ are broken according to
smaller value of $R(w)$.

### 2015-12-08 lower-bound.tex
Added proofs for mirror descent section.
Added explanations to Mirror Descent section.
Also, removed the extra words "Appendix" at several places.
Many improvements
1) Add conclusion
2) Add limits and their explanations
3) Add many bibliographic references
4) Fix formatting for most .tex files

We show a lower bound on the worst-case regret of any algorithm for OLO. The
proof is a standard probabilistic argument, which we present in
Appendix~\ref{section:lower-bound-proof}.

@@

Our upper bounds are of the form $O(\sqrt{\sup_{v \in K} f(v) \sum_{t=1}^T
\|\ell_t\|_*^2})$ where $f$ is any $1$-strongly convex function with respect to
$\|\cdot\|$.  The same upper bound is also achieved by FTRL with a constant
learning rate when the number of rounds $T$ and $\sum_{t=1}^T \|\ell_t\|_*^2$
is known upfront \cite[Chapter 2]{Shalev-Shwartz-2011}.  The lower bound is
$\Omega(D\sqrt{\sum_{t=1}^T \|\ell_t\|_*^2})$.

The gap between $D$ and $\sqrt{\sup_{v \in K} f(v)}$ can be substantial.  For
example, if $K$ is the probability simplex in $\R^d$ and $f(w) = \ln(d) +
\sum_{i=1}^d w_i \ln w_i$ is the shifted negative entropy, the
$\|\cdot\|_1$-diameter of $K$ is $2$, $f$ is non-negative and $1$-strongly
convex w.r.t. $\|\cdot\|_1$, but $\sup_{v \in K} f(v) = \ln(d)$.  On the other
hand, if the norm $\|\cdot\|_2 = \sqrt{\langle \cdot, \cdot \rangle}$ arises
from an inner product $\langle \cdot, \cdot \rangle$, the lower bound matches
the upper bounds within a constant factor.  The reason is that for any $K$ with
$\|\cdot\|_2$-diameter $D$, the function $f(w) = \frac{1}{2} \|w - w_0\|_2^2$,
where $w_0$ is an arbitrary point in $K$, is $1$-strongly convex w.r.t.
$\|\cdot\|_2$ and satisfies that $\sqrt{\sup_{v \in K} f(v)} \le D$. This leads
to the following open problem (posed also in~\cite{Kwon-Mertikopoulos-2014}):
%
\begin{quotation}
\noindent
\emph{Given a bounded convex set $K$ and a norm $\|\cdot\|$, construct a non-negative
function $f:K \to \R$ that is $1$-strongly convex with respect to $\|\cdot\|$
and minimizes $\sup_{v \in K} f(v)$.}
\end{quotation}
%
As shown in~\cite{Srebro-Sridharan-Tewari-2011}, the existence of $f$ with small
$\sup_{v \in K} f(v)$ is equivalent to the existence of an algorithm for OLO with
$\widetilde O(\sqrt{T \sup_{v \in K} f(v)})$ regret assuming $\|\ell_t\|_* \le 1$.
The $\widetilde O$ notation hides a polylogarithmic factor in $T$.

>>>>

We show a lower bound on the worst-case regret of any algorithm for OLO. The
proof is a standard probabilistic argument, which we present
in~\ref{section:lower-bound-proof}.

@@

Our upper bounds are of the form $O(\sqrt{\sup_{v \in K} f(v) \sum_{t=1}^T
\|\ell_t\|_*^2})$ where $f$ is any $1$-strongly convex function with respect to
$\|\cdot\|$.  The same upper bound is also achieved by \textsc{FTRL} with a
constant learning rate when the number of rounds $T$ and $\sum_{t=1}^T
\|\ell_t\|_*^2$ is known upfront \cite[Chapter 2]{Shalev-Shwartz-2011}.  The
lower bound is $\Omega(D\sqrt{\sum_{t=1}^T \|\ell_t\|_*^2})$.

The gap between $D$ and $\sqrt{\sup_{v \in K} f(v)}$ can be substantial.  For
example, if $K$ is the probability simplex in $\R^d$ and $f(w) = \ln(d) +
\sum_{j=1}^d w_j \ln w_j$ is the shifted negative entropy, the
$\|\cdot\|_1$-diameter of $K$ is $2$, $f$ is non-negative and $1$-strongly
convex with respect to $\|\cdot\|_1$, but $\sup_{v \in K} f(v) = \ln(d)$.  On
the other hand, if the norm $\|\cdot\|_2 = \sqrt{\langle \cdot, \cdot \rangle}$
arises from an inner product $\langle \cdot, \cdot \rangle$, the lower bound
matches the upper bounds within a constant factor.  The reason is that for any
$K$ with $\|\cdot\|_2$-diameter $D$, the function $f(w) = \frac{1}{2} \|w -
w_0\|_2^2$, where $w_0$ is an arbitrary point in $K$, is $1$-strongly convex
with respect to $\|\cdot\|_2$ and satisfies that $\sqrt{\sup_{v \in K} f(v)}
\le D$. This leads to the following open problem (posed also
in~\cite{Kwon-Mertikopoulos-2014}):
%
\begin{quotation}
\noindent
\emph{Given a bounded convex set $K$ and a norm $\|\cdot\|$, construct a
non-negative function $f:K \to \R$ that is $1$-strongly convex with respect to
$\|\cdot\|$ and minimizes $\sup_{v \in K} f(v)$.}
\end{quotation}
%
As shown in~\cite{Srebro-Sridharan-Tewari-2011}, the existence of $f$ with
small $\sup_{v \in K} f(v)$ is equivalent to the existence of an algorithm for
OLO with $\widetilde O(\sqrt{T \sup_{v \in K} f(v)})$ regret assuming
$\|\ell_t\|_* \le 1$.  The $\widetilde O$ notation hides a polylogarithmic
factor in $T$.
